Appendix S1

### 2020-06-09 analysis/paper/appendix_s1.Rmd
changed correlations to be done with Pearson correlation coefficients. Added some plots of selected metrics.

---
title: "Appendix S1"
subtitle: "Estimating Carrying Capacity for Chinook Parr using Quantile Random Forest Models"
author:
  - Kevin E. See:
      email: Kevin.See@biomark.com
      institute: [biomark]
      correspondence: true
  - Michael W. Ackerman:
      email: Mike.Ackerman@biomark.com
      institute: [biomark]
      correspondence: false
  - Richard A. Carmichael:
      email: Richie.Carmichael@biomark.com
      institute: [biomark]
      correspondence: false
  - Sarah L. Hoffmann:
      email: Sarah.Hoffmann@biomark.com
      institute: [biomark]
      correspondence: false
  - Chris Beasley:
      email: Chris.Beasley@biomark.com
      institute: [biomark]
      correspondence: false
institute:
  - biomark: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::pdf_document2:
      fig_caption: yes
      fig_height: 8.5
      fig_width: 6.5
      toc: no
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
      includes:
        in_header: ../templates/header_appendixS.tex
bibliography:
- references.bib
- packages.bib
csl: "../templates/ecological-applications.csl" # Insert path for the bib-style
---

@@

hab_corr = tibble(hab_data = list(hab_data)) %>%
  mutate(metrics = map(hab_data,
                       .f = function(x) {
                         mets = poss_hab_mets %>%
                           filter(ShortName %in% names(x)) %>%
                           pull(ShortName)
                         x %>%
                           select(one_of(mets)) %>%
                           gather(Metric, value) %>%
                           group_by(Metric) %>%
                           summarise(n_tot = n(),
                                     n_NA = sum(is.na(value)),
                                     non_0 = sum(value != 0, na.rm = T)) %>%
                           mutate(perc_NA = n_NA / n_tot,
                                  perc_non_0 = non_0 / n_tot) %>%
                           filter(perc_NA < 0.5,
                                  non_0 > 100) %>%
                           filter(!grepl('Area$', Metric),
                                  !grepl('Vol$', Metric),
                                  !Metric %in% c('Lgth_Wet', 
                                                 'Lgth_BfChnl',
                                                 'Lgth_WetChnl',
                                                 'Area_Wet', 
                                                 'Area_Bf', 
                                                 'WetVol', 
                                                 'BfVol')) %>%
                           pull(Metric)
                       }),
         cor_mat = map2(hab_data,
                        metrics,
                        .f = function(x, y) {
                          x %>%
                            select(one_of(y)) %>%
                            cor(use = 'pairwise',
                                method = 'spearman')
                        }),
         corr_mat = map(cor_mat,
                        .f = as_cordf),
         catg_cor = map2(hab_data,
                         metrics,
                         .f = function(x, y) {
                           catg_list = try(poss_hab_mets %>%
                                             filter(ShortName %in% y) %>%
                                             split(list(.$MetricCategory)) %>%
                                             map(.f = function(z) {
                                               if(nrow(z) == 1) {
                                                 return(NULL)
                                               } else {
                                                 x %>%
                                                   select(one_of(z$ShortName)) %>%
                                                   cor(use = 'pairwise',
                                                       method = 'spearman')
                                               }
                                               }))
                           return(catg_list)
                                             }))

@@

```{r corr-fig, results='hide', fig.show = "hold", fig.cap = "Correlation plots of metrics, facted by habitat category."}
hab_corr$catg_cor[[1]][!sapply(hab_corr$catg_cor[[1]], is.null)] %>%
  # map(.f = function(x) {
  #   rownames(x) <- tibble(Metric = rownames(x)) %>%
  #     left_join(hab_dict %>%
  #                 select(Metric = ShortName,
  #                        Name) %>%
  #                 distinct()) %>%
  #     pull(Name)
  #   
  #   # colnames(x) <- tibble(Metric = colnames(x)) %>%
  #   #   left_join(hab_dict %>%
  #   #               select(Metric = ShortName,
  #   #                      Name) %>%
  #   #               distinct()) %>%
  #   #   pull(Name)
  #   return(x)
  # }) %>%
    map(.f = corrplot::corrplot.mixed,
        upper = 'ellipse',
        tl.pos = 'lt',
        order = 'FPC')
```

```{r corr-fig2, results='hide', fig.show = "hold", fig.cap = "Correlation plots of metrics, facted by habitat category."}
hab_corr$catg_cor[[1]][!sapply(hab_corr$catg_cor[[1]], is.null)] %>%
  # map(.f = function(x) {
  #   # rownames(x) <- tibble(Metric = rownames(x)) %>%
  #   #   left_join(hab_dict %>%
  #   #               select(Metric = ShortName,
  #   #                      Name) %>%
  #   #               distinct()) %>%
  #   #   pull(Name)
  #   
  #   colnames(x) <- tibble(Metric = colnames(x)) %>%
  #     left_join(hab_dict %>%
  #                 select(Metric = ShortName,
  #                        Name) %>%
  #                 distinct()) %>%
  #     pull(Name)
  #   return(x)
  # }) %>%
  map(.f = ggcorrplot::ggcorrplot,
      method = 'circle',
      hc.order = T,
      type = 'lower',
      colors = c('blue4', 'white', 'red3'))
```

>>>>

---
title: "Appendix S1"
subtitle: "Estimating Carrying Capacity for Chinook Parr using Quantile Random Forest Models"
author:
  - Kevin E. See:
      email: Kevin.See@biomark.com
      institute: [biomark]
      correspondence: true
  - Michael W. Ackerman:
      email: Mike.Ackerman@biomark.com
      institute: [biomark]
      correspondence: false
  - Richard A. Carmichael:
      email: Richie.Carmichael@biomark.com
      institute: [biomark]
      correspondence: false
  - Sarah L. Hoffmann:
      email: Sarah.Hoffmann@biomark.com
      institute: [biomark]
      correspondence: false
  - Chris Beasley:
      email: Chris.Beasley@biomark.com
      institute: [biomark]
      correspondence: false
institute:
  - biomark: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      fig_height: 8.5
      fig_width: 6.5
      number_sections: FALSE
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::pdf_document2:
      fig_caption: yes
      fig_height: 8.5
      fig_width: 6.5
      toc: no
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
      includes:
        in_header: ../templates/header_appendixS.tex
bibliography:
- references.bib
- packages.bib
csl: "../templates/ecological-applications.csl" # Insert path for the bib-style
---

@@

hab_corr = tibble(hab_data = list(hab_data)) %>%
  mutate(metrics = map(hab_data,
                       .f = function(x) {
                         mets = poss_hab_mets %>%
                           filter(ShortName %in% names(x)) %>%
                           pull(ShortName)
                         x %>%
                           select(one_of(mets)) %>%
                           gather(Metric, value) %>%
                           group_by(Metric) %>%
                           summarise(n_tot = n(),
                                     n_NA = sum(is.na(value)),
                                     non_0 = sum(value != 0, na.rm = T)) %>%
                           mutate(perc_NA = n_NA / n_tot,
                                  perc_non_0 = non_0 / n_tot) %>%
                           filter(perc_NA < 0.5,
                                  non_0 > 100) %>%
                           filter(!grepl('Area$', Metric),
                                  !grepl('Vol$', Metric),
                                  !Metric %in% c('Lgth_Wet', 
                                                 'Lgth_BfChnl',
                                                 'Lgth_WetChnl',
                                                 'Area_Wet', 
                                                 'Area_Bf', 
                                                 'WetVol', 
                                                 'BfVol')) %>%
                           pull(Metric)
                       }),
         cor_mat = map2(hab_data,
                        metrics,
                        .f = function(x, y) {
                          x %>%
                            select(one_of(y)) %>%
                            cor(use = 'pairwise',
                                method = 'pearson')
                                # method = 'spearman')
                        }),
         corr_mat = map(cor_mat,
                        .f = as_cordf),
         catg_cor = map2(hab_data,
                         metrics,
                         .f = function(x, y) {
                           catg_list = try(poss_hab_mets %>%
                                             filter(ShortName %in% y) %>%
                                             split(list(.$MetricCategory)) %>%
                                             map(.f = function(z) {
                                               if(nrow(z) == 1) {
                                                 return(NULL)
                                               } else {
                                                 x %>%
                                                   select(one_of(z$ShortName)) %>%
                                                   cor(use = 'pairwise',
                                                       method = 'pearson')
                                                       # method = 'spearman')
                                               }
                                               }))
                           return(catg_list)
                                             }))

@@

```{r corr-fig, eval = F, results='hide', fig.width = 5, fig.height = 5, fig.show = "hold", fig.cap = "Correlation plots of metrics, facted by habitat category."}
hab_corr$catg_cor[[1]][!sapply(hab_corr$catg_cor[[1]], is.null)] %>%
  # map(.f = function(x) {
  #   rownames(x) <- tibble(Metric = rownames(x)) %>%
  #     left_join(hab_dict %>%
  #                 select(Metric = ShortName,
  #                        Name) %>%
  #                 distinct()) %>%
  #     pull(Name)
  #   
  #   # colnames(x) <- tibble(Metric = colnames(x)) %>%
  #   #   left_join(hab_dict %>%
  #   #               select(Metric = ShortName,
  #   #                      Name) %>%
  #   #               distinct()) %>%
  #   #   pull(Name)
  #   return(x)
  # }) %>%
    map(.f = corrplot::corrplot.mixed,
        upper = 'ellipse',
        tl.pos = 'lt',
        order = 'FPC')
```

```{r corr-fig2, results='hide', fig.width = 5, fig.height = 5, fig.show = "hold", fig.cap = "Correlation plots of metrics, facted by habitat category."}
cor_plot_df = tibble(Category = names(hab_corr$catg_cor[[1]][!sapply(hab_corr$catg_cor[[1]], is.null)]),
                     cor_mat = hab_corr$catg_cor[[1]][!sapply(hab_corr$catg_cor[[1]], is.null)]) %>%
  mutate(cor_plot = map2(cor_mat,
                         Category,
                         .f = function(x, y) {
                           ggcorrplot::ggcorrplot(x,
                                                  method = 'circle',
                                                  hc.order = T,
                                                  type = 'lower',
                                                  title = y,
                                                  colors = c('blue4', 'white', 'red3'))
                         }))

cor_plot_df$cor_plot
```

```{r selected-corr, fig.cap = 'Correlation plot of habitat metrics used in QRF model.'}
hab_data %>% 
  select(one_of(sel_hab_mets$Metric)) %>% 
  cor(use = 'pairwise',
      method = 'pearson') %>%
      # method = 'spearman') %>% 
  # as_cordf() %>%
  ggcorrplot::ggcorrplot(method = 'circle',
                         # hc.order = T,
                         type = 'lower',
                         # legend.title = "Correlation",
                         colors = c('blue4', 'white', 'red3'))
```

```{r high-cor-mets, fig.cap = 'Pairs plot of habitat metrics used in QRF model with a correlation coefficient greater than 0.5.'}
high_cor_mets = hab_data %>%
  select(one_of(sel_hab_mets$Metric)) %>%
  correlate(use = 'pairwise',
            method = 'pearson') %>%
            # method = 'spearman') %>%
  stretch(na.rm = T,
          remove.dups = T) %>%
  arrange(desc(abs(r))) %>%
  filter(abs(r) > 0.5) %>%
  select(x, y) %>%
  pivot_longer(cols = c(x, y)) %>%
  pull(value) %>%
  unique()

hab_data %>%
  select(one_of(high_cor_mets)) %>%
  GGally::ggpairs()
```
